---
title: "Brain communications- Demographics & models"
format: revealjs
editor: visual
---

## Load packages

```{r}

rm(list = ls())
pacman::p_load(tidyverse,
               janitor,
               rio,
               here,
               gt,
               GGally,
               gtExtras,
               gtsummary,
               tidymodels,
               readxl,
               dlookr,
               DataExplorer,
               arsenal,
               ggstatsplot,
               nlme,
               pROC, 
               emmeans,
               sjPlot,
               AICcmodavg,
               lme4,
               naniar,
               skimr,
               lmerTest,
               easystats,
               ggstatsplot,
               stats,
               pwr, 
               patchwork)

```

## Demographics- by patient MS

```{r}

oct_octa_clin <- import(here('data', 'processed', 'octa_oct_clin_20241201.rds'))

oct_octa_clin %>% filter(case ==  "P", lower_layer == 'GCIPL', region == 'Gen') %>%distinct(patient_id) 

# Time on DMT in days
oct_octa_clin$lastdmt_start <- as.Date(oct_octa_clin$lastdmt_start)
oct_octa_clin$exam_date <- as.Date(oct_octa_clin$exam_date)

oct_octa_clin$time_on_dmt_days <- as.numeric(oct_octa_clin$exam_date - oct_octa_clin$lastdmt_start)

oct_octa_clin$time_on_dmt_days 

# Optionally, in years
oct_octa_clin$time_on_dmt_years <- oct_octa_clin$time_on_dmt_days / 365.25
oct_octa_clin$time_on_dmt_years 

#323 patients

basic_demo_ms <- oct_octa_clin %>% filter(case ==  "P", lower_layer == 'GCIPL', region == 'Gen') %>%
  distinct(patient_id, age, sex, ethnicity, ms_type_now, disease_duration, relapses_last_2_yrs_baseline, edss, bmi, smoke_status, pack_years, units_week, hypertension, cholesterol, diabetes, lastoct_dmt, prop_dmt, het_prop_dmt, naive_at_baseline, month, dmt_1, dmt_2, dmt_3, dmt_4, dmt_5, dmt_6, dmt_1_stopped_reason, dmt_2_stopped_reason, dmt_3_stopped_reason, dmt_4_stopped_reason, dmt_5_stopped_reason, lastoct_dmt, time_on_dmt_years) 

#Age
summary(basic_demo_ms$age) 
sd(basic_demo_ms$age)
hist(basic_demo_ms$age, col = "firebrick3", breaks=50)
 
#Gender
table(basic_demo_ms$sex)
 
#Ethnicity
table(basic_demo_ms$ethnicity)
 
#Phenotype
table(basic_demo_ms$ms_type_now)
 
#Disease duration
summary(basic_demo_ms$disease_duration)
sd(basic_demo_ms$disease_duration)
hist(basic_demo_ms$disease_duration, col = "firebrick3", breaks=50)

basic_demo_ms <- basic_demo_ms%>%mutate(bl_arr= relapses_last_2_yrs_baseline/2)%>%relocate(bl_arr, .after= relapses_last_2_yrs_baseline)

#Baseline ARR
summary(basic_demo_ms$bl_arr)
sd(basic_demo_ms$bl_arr)
hist(basic_demo_ms$bl_arr, col = "firebrick3", breaks=50)

basic_demo_ms$edss <- as.numeric(basic_demo_ms$edss)

#EDSS
class(basic_demo_ms$edss)
summary(basic_demo_ms$edss, na.rm = T)
sd(basic_demo_ms$edss, na.rm = T)
hist(basic_demo_ms$edss, col = "firebrick3", breaks=50) 

#Current DMT
table(basic_demo_ms$lastoct_dmt)

basic_demo_ms %>%filter(is.na(lastoct_dmt))

#DMT proportion
summary(basic_demo_ms$prop_dmt, na.rm = T)
sd(basic_demo_ms$prop_dmt, na.rm = T)
hist(basic_demo_ms$prop_dmt, col = "firebrick3", breaks=50)

#HET
summary(basic_demo_ms$het_prop_dmt, na.rm = T)
sd(basic_demo_ms$het_prop_dmt, na.rm = T)
hist(basic_demo_ms$het_prop_dmt, col = "firebrick3", breaks=50)

#Time on last DMT
summary(basic_demo_ms$time_on_dmt_years, na.rm = T)
sd(basic_demo_ms$time_on_dmt_years, na.rm = T)
hist(basic_demo_ms$time_on_dmt_years, col = "firebrick3", breaks=50)

basic_demo_ms%>%
  count(lastoct_dmt)

#Smoke status
table(basic_demo_ms$smoke_status)

#pack-years
#Need to filter out 0 pack years
summary(basic_demo_ms$pack_years,na.rm = T)
sd(basic_demo_ms$pack_years,na.rm = T)
hist(basic_demo_ms$pack_years, col = "firebrick3", breaks=50)

#BMI
summary(basic_demo_ms$bmi,na.rm = T)
sd(basic_demo_ms$bmi,na.rm = T)
hist(basic_demo_ms$bmi, col = "firebrick3", breaks=50)

#Alcohol
basic_demo_ms$units_week <- as.numeric(basic_demo_ms$units_week)
oct_octa_clin$units_week <- as.numeric(oct_octa_clin$units_week)

summary(basic_demo_ms$units_week,na.rm = T)
sd(basic_demo_ms$units_week)
hist(basic_demo_ms$units_week, col = "firebrick3", breaks=50)

#Hypertension
table(basic_demo_ms$hypertension)

basic_demo_ms%>%filter(hypertension == "yes")

#Cholesterol
table(basic_demo_ms$cholesterol)

#Diabetes
table(basic_demo_ms$diabetes)

#Naive
table(basic_demo_ms$naive_at_baseline, basic_demo_ms$month)
#96 truly naive
#184 on 1st DMT

#Switch reason
basic_demo_ms_not_naive<-basic_demo_ms%>%filter(naive_at_baseline=="no")
basic_demo_ms_naive<-basic_demo_ms%>%filter(naive_at_baseline=="yes")

basic_demo_ms_not_naive%>%select(patient_id,dmt_1_stopped_reason, dmt_2_stopped_reason, dmt_3_stopped_reason, dmt_4_stopped_reason, dmt_5_stopped_reason)

basic_demo_ms_naive%>%tabyl(lastoct_dmt)

#filter last DMT
basic_demo_ms_not_naive <- basic_demo_ms_not_naive %>%
  mutate(
    lastoct_dmt = case_when(
      !is.na(dmt_6) ~ dmt_6,
      is.na(dmt_6) & !is.na(dmt_5) ~ dmt_5,
      is.na(dmt_5) & !is.na(dmt_4) ~ dmt_4,
      is.na(dmt_4) & !is.na(dmt_3) ~ dmt_3,
      is.na(dmt_3) & !is.na(dmt_2) ~ dmt_2,
      is.na(dmt_2) & !is.na(dmt_1) ~ dmt_1,
      .default = NA
    ))

basic_demo_ms_not_naive<-basic_demo_ms_not_naive %>%
  mutate(
    last_dmt_stop_reason = pmap_chr(
      list(dmt_1_stopped_reason, dmt_2_stopped_reason, dmt_3_stopped_reason,
           dmt_4_stopped_reason, dmt_5_stopped_reason),
      ~ {
        reasons <- c(...)
        # Take the last non-NA stop reason
        last_reason <- rev(reasons[!is.na(reasons)])[1]
        ifelse(is.null(last_reason), NA_character_, last_reason)
      }
    )
  )

basic_demo_ms_not_naive %>%
  count(last_dmt_stop_reason, sort = TRUE)


#Type of switch
basic_demo_ms_not_naive <- basic_demo_ms_not_naive %>%
  mutate(
    # Find the current/last DMT
    lastoct_dmt = case_when(
      !is.na(dmt_6) ~ dmt_6,
      is.na(dmt_6) & !is.na(dmt_5) ~ dmt_5,
      is.na(dmt_5) & !is.na(dmt_4) ~ dmt_4,
      is.na(dmt_4) & !is.na(dmt_3) ~ dmt_3,
      is.na(dmt_3) & !is.na(dmt_2) ~ dmt_2,
      is.na(dmt_2) & !is.na(dmt_1) ~ dmt_1,
      TRUE ~ NA_character_
    ),
    # Find the previous DMT (the one right before last)
    prev_dmt = pmap_chr(
      list(dmt_1, dmt_2, dmt_3, dmt_4, dmt_5, dmt_6),
      ~ {
        dmts <- c(...)
        # Keep only non-NA DMTs
        dmts_non_na <- dmts[!is.na(dmts)]
        # Take the second-to-last DMT (NA if only one)
        if(length(dmts_non_na) < 2) NA_character_ else dmts_non_na[length(dmts_non_na)-1]
      }
    ),
    # Combine to show switch
    last_switch = ifelse(!is.na(prev_dmt) & !is.na(lastoct_dmt),
                         paste(prev_dmt, "→", lastoct_dmt),
                         NA_character_)
  )

basic_demo_ms_not_naive %>%
  filter(!is.na(last_switch)) %>%
  count(last_switch) %>%
  mutate(percent = n / sum(n) * 100)

```

## Demographics- HC

```{r}

oct_octa_clin %>% filter(case ==  "C", lower_layer == 'GCIPL', region == 'Gen')%>%distinct(patient_id) 

#80 healthy controls

basic_demo_hc <- oct_octa_clin %>% filter(case ==  "C", lower_layer == 'GCIPL', region == 'Gen') %>%
  distinct(patient_id, age, sex, smoke_status, pack_years, units_week, hypertension, cholesterol, diabetes)

#Age
 summary(basic_demo_hc$age)
 sd(basic_demo_hc$age)
 hist(basic_demo_hc$age, col = "firebrick3", breaks=50)

#Gender
 table(basic_demo_hc$sex)
 
#Smoke status
table(basic_demo_hc$smoke_status)

summary(basic_demo_hc$pack_years,na.rm = T)
sd(basic_demo_hc$pack_years,na.rm = T)
hist(basic_demo_hc$pack_years, col = "firebrick3", breaks=50)

#Alcohol
summary(basic_demo_hc$units_week,na.rm = T)
sd(basic_demo_hc$units_week,na.rm = T)
hist(basic_demo_hc$units_week, col = "firebrick3", breaks=50)

#Hypertension
table(basic_demo_hc$hypertension)

#Cholesterol
table(basic_demo_hc$cholesterol)

#Diabetes
table(basic_demo_hc$diabetes)

```

## Check summary based on eyes

```{r}

general <- oct_octa_clin %>% 
filter(region == 'Gen' & lower_layer == 'GCIPL') 

general%>%distinct(patient_id)

general%>%filter(!(is.na(vad)))%>%distinct(patient_id)

general_P <- general %>%
  filter(case == 'P')

general_P%>%distinct(patient_id)

general_P %>% count(on)

general_P %>% 
  count(oct_optic_neuritis)

```

## t-tests

```{r}

general_demo <- general%>%distinct(patient_id, age, sex, case, smoke_status, pack_years, units_week, hypertension, cholesterol, diabetes)

general_demo <-general_demo %>% mutate(cardiovascular = case_when(
  hypertension== "yes"|cholesterol == "yes"|diabetes =="yes" ~ "yes",
  .default = "no"))


age.ttest <- t.test(basic_demo_hc$age, basic_demo_ms$age, var.equal = TRUE)
age.ttest

#no significant difference

#Gender differences- HC & MS

ggpiestats(
  data = general_demo,
  x= sex,
  y=case
) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(
    colour= "black",
    size= 14,
    hjust=0
  )
)

#Smokers

ggpiestats(
  data = general_demo,
  x= smoke_status,
  y=case
) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(
    colour= "black",
    size= 14,
    hjust=0
  )
)

#CV risk factors

ggpiestats(
  data = general_demo,
  x= cardiovascular,
  y=case
) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(
    colour= "black",
    size= 14,
    hjust=0
  )
)

```

## OCT-summary

```{r}

#Patients
summary(general_P$av_6mm)
sd(general_P$av_6mm)
hist(general_P$av_6mm, col = "firebrick3", breaks=50)

summary(general_P$rnfl_mean_g)
sd(general_P$rnfl_mean_g,na.rm = T)
hist(general_P$rnfl_mean_g, col = "firebrick3", breaks=50)

#NON- OCT

summary(general_non$av_6mm)
sd(general_non$av_6mm)
hist(general_non$av_6mm, col = "firebrick3", breaks=50)

summary(general_non$rnfl_mean_g)
sd(general_non$rnfl_mean_g)
hist(general_non$rnfl_mean_g, col = "firebrick3", breaks=50)


#ON- OCT
summary(general_on$av_6mm)
sd(general_on$av_6mm)
hist(general_on$av_6mm, col = "firebrick3", breaks=50)

summary(general_on$rnfl_mean_g)
sd(general_on$rnfl_mean_g, na.rm = T)
hist(general_on$rnfl_mean_g, col = "firebrick3", breaks=50)


#HCs
summary(general_C$av_6mm)
sd(general_C$av_6mm)
hist(general_C$av_6mm, col = "firebrick3", breaks=50)

summary(general_C$rnfl_mean_g)
sd(general_C$rnfl_mean_g)
hist(general_C$rnfl_mean_g, col = "firebrick3", breaks=50)


```

## Regression models -eye status 

```{r}

#GCIPL
case_gcipl_lmm <- lme(av_6mm ~ eye_status2 + age + sex, data = general, random=~1|patient_id, na.action=na.omit)
summary(case_gcipl_lmm)
intervals(case_gcipl_lmm)
multicollinearity(case_gcipl_lmm)

emmeans_gcipl <-emmeans(case_gcipl_lmm, "eye_status2")
emmeans(emmeans_gcipl, pairwise ~ "eye_status2", infer = T) 
emmeans(emmeans_gcipl, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_gcipl)
plot_model(case_gcipl_lmm, type = 'emm', terms = c('eye_status2'), show.values = TRUE)

boxplot(av_6mm ~ eye_status2, data=general, main="Macular GCIPL thickness in eyes", 
   xlab="Eye status", ylab="mGCIPL (microns)", names= c("HC", "NON", "ON"), show.values = TRUE)

ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = av_6mm,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "GCIPL based on Eye status",
    x = "Eye Status",
    y = "macular GCIPL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

gcipl_box <- ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = av_6mm,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "GCIPL based on Eye status",
    x = "Eye Status",
    y = "macular GCIPL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")



#pRNFL
case_prnfl_lmm <- lme(rnfl_mean_g ~ eye_status2 + age + sex, data = general, random=~1|patient_id, na.action=na.omit)
summary(case_prnfl_lmm)
intervals(case_prnfl_lmm)
multicollinearity(case_prnfl_lmm)

emmeans_rnfl <-emmeans(case_prnfl_lmm, "eye_status2")
emmeans(emmeans_rnfl, pairwise ~ "eye_status2", infer = T) 
emmeans(emmeans_rnfl, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_rnfl)
plot_model(case_prnfl_lmm, type = 'emm', terms = c('eye_status2'), show.values = TRUE)

boxplot(rnfl_mean_g ~ eye_status2, data=general, main="pRNFL thickness in eyes", 
   xlab="Eye status", ylab="pRNFL (microns)", names= c("HC", "NON", "ON"), show.values = TRUE)

ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = rnfl_mean_g,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "pRNFL based on Eye status",
    x = "Eye Status",
    y = "pRNFL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

rnfl_box<-ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = rnfl_mean_g,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "pRNFL based on Eye status",
    x = "Eye Status",
    y = "pRNFL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")


#INL
case_inl_lmm <- lme(av_6mm ~ eye_status2 + age + sex, data = general_inl, random=~1|patient_id, na.action=na.omit)
summary(case_inl_lmm)
intervals(case_inl_lmm)
multicollinearity(case_inl_lmm)

emmeans_inl <-emmeans(case_inl_lmm, "eye_status2")
emmeans(emmeans_inl, pairwise ~ "eye_status2", infer = T) 
emmeans(emmeans_inl, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_inl)
plot_model(case_inl_lmm, type = 'emm', terms = c('eye_status2'), show.values = TRUE)

boxplot(av_6mm  ~ eye_status2, data=general_inl, main="Macular INL thickness in eyes", 
   xlab="Eye status", ylab="INL (microns)", names= c("HC", "NON", "ON"), show.values = TRUE)

ggplot(general_inl, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = av_6mm,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "INL based on Eye status",
    x = "Eye Status",
    y = "INL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

inl_box <- ggplot(general_inl, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = av_6mm,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "INL based on Eye status",
    x = "Eye Status",
    y = "INL (microns)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

#vad
case_vad_lmm <- lme(vad ~ eye_status2 + age + sex, data = general, random=~1|patient_id, na.action=na.omit)
summary(case_vad_lmm)
intervals(case_vad_lmm)
multicollinearity(case_vad_lmm)

emmeans_vad <-emmeans(case_vad_lmm, "eye_status2")
emmeans(emmeans_vad, pairwise ~ "eye_status2", infer = T) 
emmeans(emmeans_vad, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_inl)
plot_model(case_vad_lmm, type = 'emm', terms = c('eye_status2'), show.values = TRUE)

boxplot(vad ~ eye_status2, data=general, main="Vessel area density in eyes", 
   xlab="Eye status", ylab="Vessel area density (%)", names= c("HC", "NON", "ON"), show.values = TRUE)

ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = vad,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "VAD based on Eye status",
    x = "Eye Status",
    y = "Vessel area density (%)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

vad_box<-ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = vad,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "VAD based on Eye status",
    x = "Eye Status",
    y = "Vessel area density (%)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")


#vld
case_vld_lmm <- lme(vld_pct ~ eye_status2 + age + sex, data = general, random=~1|patient_id, na.action=na.omit)
summary(case_vld_lmm)
intervals(case_vld_lmm)
multicollinearity(case_vld_lmm)

emmeans_vld <-emmeans(case_vld_lmm, "eye_status2")
emmeans(emmeans_vld, pairwise ~ "eye_status2", infer = T) 
emmeans(emmeans_vld, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_vld)
plot_model(case_vld_lmm, type = 'emm', terms = c('eye_status2'), show.values = TRUE)


ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = vld_pct,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "VLD based on Eye status",
    x = "Eye Status",
    y = "Vessel length density in pixels (mm-1)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")


vld_box<-ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = vld_pct,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "VLD based on Eye status",
    x = "Eye Status",
    y = "Vessel length density in pixels (mm-1)"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

#fd
case_fd <- lme(fd ~ eye_status2 + age + sex, data = general, random=~1|patient_id, na.action=na.omit)
summary(case_fd)
intervals(case_fd)
multicollinearity(case_fd)

emmeans_fd <-emmeans(case_fd, "eye_status2")
emmeans(emmeans_fd, pairwise ~ "eye_status2", infer = T)  
emmeans(emmeans_fd, pairwise ~ "eye_status2", infer = T) %>% plot(comparisons = T)
tidy(emmeans_fd)
plot_model(case_fd, type = 'emm', terms = c('eye_status2'), show.values = TRUE)

boxplot(fd ~ eye_status2,data=general, main="Fractal dimension in eyes", 
   xlab="Eye status", ylab="Fractal dimension", names= c("HC", "NON", "ON"))

ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = fd,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "FD based on Eye status",
    x = "Eye Status",
    y = "Fractal dimension"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

fd_box<-ggplot(general, aes(
        x = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON")),
        y = fd,
        fill = factor(eye_status2, levels = c(0, 1, 2), labels = c("HC", "NON", "ON"))
      )) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
  labs(
    title = "FD based on Eye status",
    x = "Eye Status",
    y = "Fractal dimension"
  ) +
  scale_fill_manual(values = c("skyblue", "orange", "pink")) +
  theme_minimal() +
  theme(legend.position = "none")

```

## ROC curves-averaging NON eyes

```{r}

# Average metrics for bilateral NON eyes
bilat_non <- general %>%
  filter(eye_status2 == "1") %>%
  group_by(patient_id) %>%
  filter(n() == 2) %>%  # only patients with both eyes
  summarise(
    vad_avg = mean(vad),
    vld_avg = mean(vld_pct),
    fd_avg = mean(fd),
    gcipl_avg = mean(av_6mm),
    rnfl_avg = mean(rnfl_mean_g),
    age = first(age),
    sex = first(sex),
    eye_status2 = first(eye_status2)  # all NON
  )

# Average metrics for bilateral HCs
bilat_hc <- general %>%
  filter(eye_status2 == "0") %>%
  group_by(patient_id) %>%
  filter(n() == 2) %>%
  summarise(
    vad_avg = mean(vad),
    vld_avg = mean(vld_pct),
    fd_avg = mean(fd),
    gcipl_avg = mean(av_6mm),
    rnfl_avg = mean(rnfl_mean_g),
    age = first(age),
    sex = first(sex),
    eye_status2 = first(eye_status2)  # all HC
  )

# Combine datasets
roc_data <- bind_rows(bilat_non, bilat_hc)

roc_data$eye_status2<-as.numeric(roc_data$eye_status2)

roc_vad <- roc(roc_data$eye_status2, roc_data$vad_avg)
plot(roc_vad, col="blue", lwd=2, main="ROC Curve: VAD (NON vs HC)")
pROC::auc(roc_vad)
cat("AUC VAD:", pROC::auc(roc_vad), "\n")
#AUC= 0.6295

#Adjust for age and sex
#VAD
model_vad_auc <- glm(eye_status2 ~ vad_avg + age + sex, 
                 family = binomial, data = roc_data)

# Get predicted probabilities
roc_data$pred_vad <- predict(model_vad_auc, type = "response")

# ROC and AUC
roc_vad2 <- roc(roc_data$eye_status2, roc_data$pred_vad)
plot(roc_vad2, col="blue", lwd=2, main="ROC Curve: VAD (adjusted)")
pROC::auc(roc_vad2)
ci.auc(roc_vad2)
#AUC 0.6405

#VLD
model_vld_auc <- glm(eye_status2 ~ vld_avg + age + sex, 
                 family = binomial, data = roc_data)
# Get predicted probabilities
roc_data$pred_vld <- predict(model_vld_auc, type = "response")

# ROC and AUC
roc_vld <- roc(roc_data$eye_status2, roc_data$pred_vld)
plot(roc_vld, col="red", lwd=2, main="ROC Curve: VLD (adjusted)")
pROC::auc(roc_vld)
ci.auc(roc_vld)
#AUC 0.6459

#FD
model_fd_auc <- glm(eye_status2 ~ fd_avg + age + sex, 
                 family = binomial, data = roc_data)
# Get predicted probabilities
roc_data$pred_fd <- predict(model_fd_auc, type = "response")

# ROC and AUC
roc_fd <- roc(roc_data$eye_status2, roc_data$pred_fd)
plot(roc_fd, col="green", lwd=2, main="ROC Curve: FD (adjusted)")
pROC::auc(roc_fd)
ci.auc(roc_fd)
#AUC 0.6458

#GCIPL
model_gcipl_auc <- glm(eye_status2 ~ gcipl_avg + age + sex, 
                 family = binomial, data = roc_data)
# Get predicted probabilities
roc_data$pred_gcipl <- predict(model_gcipl_auc, type = "response")

# ROC and AUC
roc_gcipl <- roc(roc_data$eye_status2, roc_data$pred_gcipl)
plot(roc_gcipl, col="purple", lwd=2, main="ROC Curve:mGCIPL (adjusted)")
pROC::auc(roc_gcipl)
ci.auc(roc_gcipl)
#AUC 0.6725

#pRNFL
model_rnfl_auc <- glm(eye_status2 ~ rnfl_avg + age + sex, 
                 family = binomial, data = roc_data)
# Get predicted probabilities
roc_data$pred_rnfl <- predict(model_rnfl_auc, type = "response")

# ROC and AUC
roc_rnfl <- roc(roc_data$eye_status2, roc_data$pred_rnfl)
plot(roc_rnfl, col="orange", lwd=2, main="ROC Curve:pRNFL (adjusted)")
pROC::auc(roc_rnfl)
ci.auc(roc_rnfl)
#AUC 0.6258

#Compare using Delong`s test
roc.test(roc_gcipl, roc_vad, method="delong")
roc.test(roc_rnfl, roc_vad, method="delong")
roc.test(roc_rnfl, roc_gcipl, method="delong")
roc.test(roc_vld, roc_gcipl, method="delong")
roc.test(roc_fd, roc_gcipl, method="delong")
roc.test(roc_fd, roc_rnfl, method="delong")
roc.test(roc_vld, roc_rnfl, method="delong")

```

## EDSS- interactions with disease duration and ON

```{r}

#gcipl
model_edss_gcipl<- lme(av_6mm ~ edss*disease_duration + oct_optic_neuritis + age + sex + prop_dmt , data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_gcipl)
intervals(model_edss_gcipl)
plot_model(model_edss_gcipl, type= "emm", terms= c("edss", "disease_duration"), labs(y= "GCIPL", x = "EDSS"))
plot_model(model_edss_gcipl,
  type = "emm",
  terms = c("edss", "disease_duration"),
  labs = list(y = "GCIPL", x = "EDSS"),
  colors = c("#0072B2", "#E69F00", "#CC79A7"))

edss_gcipl<-plot_model(model_edss_gcipl,
  type = "emm",
  terms = c("edss", "disease_duration"),
  labs = list(y = "GCIPL", x = "EDSS"),
  colors = c("#0072B2", "#E69F00", "#CC79A7"))

multicollinearity(model_edss_gcipl)
emtrends(model_edss_gcipl, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)

model_edss_gcipl2<- lme(av_6mm ~ edss*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_gcipl2)
intervals(model_edss_gcipl2)
plot_model(model_edss_gcipl2, type= "emm", terms= c("edss", "oct_optic_neuritis"))
multicollinearity(model_edss_gcipl2)
emtrends(model_edss_gcipl2, pairwise ~ oct_optic_neuritis, var = "edss", infer = T)

anova(model_edss_gcipl, model_edss_gcipl2) 
compare_models(model_edss_gcipl, model_edss_gcipl2) 
plot(compare_performance(model_edss_gcipl, model_edss_gcipl2, rank = TRUE, verbose = FALSE))

#prnfl
model_edss_prnfl<- lme(rnfl_mean_g ~ edss*disease_duration + age + sex +  oct_optic_neuritis + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_prnfl) 
intervals(model_edss_prnfl)
multicollinearity(model_edss_prnfl)
emtrends(model_edss_prnfl, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)
plot_model(model_edss_prnfl, type= "emm", terms= c("edss", "disease_duration"))

model_edss_prnfl2<- lme(rnfl_mean_g ~ edss*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_prnfl2)
intervals(model_edss_prnfl2)
plot_model(model_edss_prnfl2, type= "emm", terms= c("edss", "oct_optic_neuritis"))
multicollinearity(model_edss_prnfl2)
emtrends(model_edss_prnfl2, pairwise ~ oct_optic_neuritis, var = "edss", infer = T)

anova(model_edss_prnfl, model_edss_prnfl2) 
compare_models(model_edss_prnfl, model_edss_prnfl2) 
plot(compare_performance(model_edss_prnfl, model_edss_prnfl2, rank = TRUE, verbose = FALSE))


#vad
model_edss_vad<- lme(vad ~ edss*disease_duration + age + sex + oct_optic_neuritis + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
multicollinearity(model_edss_vad)
summary(model_edss_vad)
intervals(model_edss_vad)
plot_model(model_edss_vad, type= "emm", terms= c("edss", "disease_duration"))
plot_model(model_edss_vad,
  type = "emm",
  terms = c("edss", "disease_duration"),
  labs = list(y = "vad", x = "EDSS"),
  colors = c("#0072B2", "#E69F00", "#CC79A7"))

edss_vad<-plot_model(model_edss_vad,
  type = "emm",
  terms = c("edss", "disease_duration"),
  labs = list(y = "vad", x = "EDSS"),
  colors = c("#0072B2", "#E69F00", "#CC79A7"))

emtrends(model_edss_vad, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)

ggplot(general_P, aes(x=edss, y=vad)) + geom_point() + geom_smooth(method="lm") + ggtitle("EDSS vs VAD")

model_edss_vad2<- lme(vad ~ edss*oct_optic_neuritis + disease_duration + age + sex + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
multicollinearity(model_edss_vad2)
summary(model_edss_vad2)
intervals(model_edss_vad2)
plot_model(model_edss_vad2, type= "emm", terms= c("edss", "oct_optic_neuritis"))
emtrends(model_edss_vad2, pairwise ~ oct_optic_neuritis, var = "edss", infer = T)

anova(model_edss_vad, model_edss_vad2) 
compare_models(model_edss_vad, model_edss_vad2) 
plot(compare_performance(model_edss_vad, model_edss_vad2, rank = TRUE, verbose = FALSE))

#vld
model_edss_vld<- lme(vld_pct ~ edss*disease_duration + age + sex + oct_optic_neuritis + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_vld)
intervals(model_edss_vld)
multicollinearity(model_edss_vld)

plot_model(model_edss_vld, type= "emm", terms= c("edss", "disease_duration"))
emtrends(model_edss_vld, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 12, 14)), var = "edss", infer = T)

model_edss_vld2<- lme(vld_pct ~ edss*oct_optic_neuritis + disease_duration + age + sex + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
multicollinearity(model_edss_vld2)
summary(model_edss_vld2)
intervals(model_edss_vld2)
plot_model(model_edss_vld2, type= "emm", terms= c("edss", "oct_optic_neuritis"))
emtrends(model_edss_vld2, pairwise ~ oct_optic_neuritis, var = "edss", infer = T)

anova(model_edss_vld, model_edss_vld2) 
compare_models(model_edss_vld, model_edss_vld2) 
plot(compare_performance(model_edss_vld, model_edss_vld2, rank = TRUE, verbose = FALSE))

#FD
model_edss_fd<- lme(fd ~ edss*disease_duration + age + sex + oct_optic_neuritis +  prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_fd)
intervals(model_edss_fd)
multicollinearity(model_edss_fd)

plot_model(model_edss_fd, type= "emm", terms= c("edss", "disease_duration"))
emtrends(model_edss_fd, pairwise ~ disease_duration, at =list(disease_duration = c(1, 5, 6, 7, 12, 14)), var = "edss", infer = T)

model_edss_fd2<- lme(fd ~ edss*oct_optic_neuritis + disease_duration + age + sex + prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_fd2) 
intervals(model_edss_fd2)
multicollinearity(model_edss_fd2)

plot_model(model_edss_fd2, type= "emm", terms= c("edss", "oct_optic_neuritis"))
emtrends(model_edss_fd2, pairwise ~ oct_optic_neuritis, var = "edss", infer = T)

```

## Sensitivity analysis- % on HE-DMT

```{r}

model_edss_vad_het<- lme(vad ~ edss*disease_duration + age + sex + oct_optic_neuritis + het_prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
multicollinearity(model_edss_vad_het)
summary(model_edss_vad_het)
intervals(model_edss_vad_het)
plot_model(model_edss_vad_het, type= "emm", terms= c("edss", "disease_duration"))
emtrends(model_edss_vad_het, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)

model_edss_gcipl_het<- lme(av_6mm ~ edss*disease_duration + oct_optic_neuritis + age + sex + het_prop_dmt , data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_gcipl_het)
intervals(model_edss_gcipl_het)
plot_model(model_edss_gcipl_het, type= "emm", terms= c("edss", "disease_duration"), labs(y= "GCIPL", x = "EDSS"))
multicollinearity(model_edss_gcipl_het)
emtrends(model_edss_gcipl_het, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)

model_edss_prnfl_het<- lme(rnfl_mean_g ~ edss*disease_duration + age + sex +  oct_optic_neuritis + het_prop_dmt, data = general_P, random=~1|patient_id, na.action=na.omit)
summary(model_edss_prnfl_het) 
intervals(model_edss_prnfl_het)
multicollinearity(model_edss_prnfl_het)
emtrends(model_edss_prnfl_het, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "edss", infer = T)
plot_model(model_edss_prnfl_het, type= "emm", terms= c("edss", "disease_duration"))

```

## Visual outcomes- summary

```{r}


general%>%filter(is.na(logmar_min))%>%filter(case=="P")

#Logmar
summary(general_P$logmar_min, na.rm = T)
sd(general_P$logmar_min, na.rm = T)
hist(general_P$logmar_min, col = "firebrick3", breaks=50)

#Sloan 2.5%

summary(general_P$sloan_cont_2_5, na.rm = T)
sd(general_P$sloan_cont_2_5, na.rm = T)
hist(general_P$sloan_cont_2_5, col = "firebrick3", breaks=50)

#Sloan 1.25%

summary(general_P$sloan_cont_1_25, na.rm = T)
sd(general_P$sloan_cont_1_25, na.rm = T)
hist(general_P$sloan_cont_1_25, col = "firebrick3", breaks=50)

#Ishihara

summary(general_P$lcva_ishihara_plates, na.rm = T)
sd(general_P$lcva_ishihara_plates, na.rm = T)
hist(general_P$lcva_ishihara_plates, col = "firebrick3", breaks=50)


#NON eyes

#Logmar
summary(general_non$logmar_min, na.rm = T)
sd(general_non$logmar_min, na.rm = T)
hist(general_non$logmar_min, col = "firebrick3", breaks=50)

#Sloan 2.5%

summary(general_non$sloan_cont_2_5, na.rm = T)
sd(general_non$sloan_cont_2_5, na.rm = T)
hist(general_non$sloan_cont_2_5, col = "firebrick3", breaks=50)

#Sloan 1.25%

summary(general_non$sloan_cont_1_25, na.rm = T)
sd(general_non$sloan_cont_1_25, na.rm = T)
hist(general_non$sloan_cont_1_25, col = "firebrick3", breaks=50)

#Ishihara

summary(general_non$lcva_ishihara_plates, na.rm = T)
sd(general_non$lcva_ishihara_plates, na.rm = T)
hist(general_non$lcva_ishihara_plates, col = "firebrick3", breaks=50)

#ON eyes

#Logmar
summary(general_on$logmar_min, na.rm = T)
sd(general_on$logmar_min, na.rm = T)
hist(general_on$logmar_min, col = "firebrick3", breaks=50)

#Sloan 2.5%

summary(general_on$sloan_cont_2_5, na.rm = T)
sd(general_on$sloan_cont_2_5, na.rm = T)
hist(general_on$sloan_cont_2_5, col = "firebrick3", breaks=50)

#Sloan 1.25%

summary(general_on$sloan_cont_1_25, na.rm = T)
sd(general_on$sloan_cont_1_25, na.rm = T)
hist(general_on$sloan_cont_1_25, col = "firebrick3", breaks=50)

#Ishihara

summary(general_on$lcva_ishihara_plates, na.rm = T)
sd(general_on$lcva_ishihara_plates, na.rm = T)
hist(general_on$lcva_ishihara_plates, col = "firebrick3", breaks=50)


```

## Visual outcomes- Logmar models 

```{r}

logmar_gcipl2 <-  lme(logmar_min~ av_6mm*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))} , random=~1|patient_id, na.action=na.omit)
multicollinearity(logmar_gcipl2)
summary(logmar_gcipl2)
intervals(logmar_gcipl2)

plot_model(logmar_gcipl2, type= "emm", terms= c("av_6mm", "oct_optic_neuritis"), show.data = T)
logmar_gcipl<-plot_model(logmar_gcipl2, type= "emm", terms= c("av_6mm", "oct_optic_neuritis"), show.data = T)
emtrends(logmar_gcipl2, pairwise ~ oct_optic_neuritis, var = "av_6mm", infer = T)

logmar_gcipl3 <-  lme(logmar_min ~ av_6mm*disease_duration + oct_optic_neuritis + sex + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_gcipl3)
intervals(logmar_gcipl3)
check_model(logmar_gcipl3)
multicollinearity(logmar_gcipl3)

plot_model(logmar_gcipl3, type= "emm", terms= c("av_6mm", "disease_duration"))
emtrends(logmar_gcipl3, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "av_6mm", infer = T)

anova(logmar_gcipl2, logmar_gcipl3) 
compare_models(logmar_gcipl2, logmar_gcipl3) 
plot(compare_performance(logmar_gcipl2, logmar_gcipl3, rank = TRUE, verbose = FALSE))

logmar_rnfl2 <-  lme(logmar_min ~ rnfl_mean_g*oct_optic_neuritis + age + sex  + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(logmar_rnfl2)
summary(logmar_rnfl2) 
intervals(logmar_rnfl2)

plot_model(logmar_rnfl2, type= "emm", terms= c("rnfl_mean_g", "oct_optic_neuritis"))
logmar_rnfl<-plot_model(logmar_rnfl2, type= "emm", terms= c("rnfl_mean_g", "oct_optic_neuritis"))
emtrends(logmar_rnfl2, pairwise ~ oct_optic_neuritis, var = "rnfl_mean_g", infer = T)

logmar_rnfl3 <-  lme(logmar_min ~ rnfl_mean_g*disease_duration + age + sex  + oct_optic_neuritis + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(logmar_rnfl3)
summary(logmar_rnfl3) 
intervals(logmar_rnfl3)

plot_model(logmar_rnfl3, type= "emm", terms= c("rnfl_mean_g", "disease_duration"))
logmar_rnfl<-plot_model(logmar_rnfl3, type= "emm", terms= c("rnfl_mean_g", "disease_duration"))
emtrends(logmar_rnfl3, pairwise ~ disease_duration, var = "rnfl_mean_g", infer = T)

logmar_vad2 <-  lme(logmar_min ~ vad*oct_optic_neuritis + sex  + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_vad2)
intervals(logmar_vad2)
check_model(logmar_vad2)
multicollinearity(logmar_vad2)

plot_model(logmar_vad2, type= "emm", terms= c("vad", "oct_optic_neuritis"), show.data=T)
logmar_vad<-plot_model(logmar_vad2, type= "emm", terms= c("vad", "oct_optic_neuritis"), show.data=T)
emtrends(logmar_vad2, pairwise ~ oct_optic_neuritis, var = "vad", infer = T)

logmar_vad3 <-  lme(logmar_min ~ vad*disease_duration + oct_optic_neuritis + sex + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_vad3)
intervals(logmar_vad3)
check_model(logmar_vad3)
multicollinearity(logmar_vad3)

plot_model(logmar_vad3, type= "emm", terms= c("vad", "disease_duration"))
emtrends(logmar_vad3, pairwise ~ disease_duration, at =list(disease_duration = c(1, 7, 9, 10, 12, 14)), var = "vad", infer = T)

anova(logmar_vad2, logmar_vad3) 
compare_models(logmar_vad2, logmar_vad3) 
plot(compare_performance(logmar_vad2, logmar_vad3, rank = TRUE, verbose = FALSE))

logmar_vld2 <-  lme(logmar_min ~ vld_pct*oct_optic_neuritis + age + sex  + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_vld2)
intervals(logmar_vld2)
multicollinearity(logmar_vld2)

plot_model(logmar_vld2, type= "emm", terms= c("vld_pct", "oct_optic_neuritis"))
emtrends(logmar_vld2, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T)

logmar_vld3 <-  lme(logmar_min ~ vld_pct*disease_duration + oct_optic_neuritis age + sex  + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_vld3)
intervals(logmar_vld3)
multicollinearity(logmar_vld3)

plot_model(logmar_vld3, type= "emm", terms= c("vld_pct", "disease_duration"))
emtrends(logmar_vld3, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T)

logmar_fd2 <-  lme(logmar_min ~ fd*oct_optic_neuritis + age + sex  + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_fd2)
intervals(logmar_fd2)
multicollinearity(logmar_fd2)

plot_model(logmar_fd2, type= "emm", terms= c("fd", "oct_optic_neuritis"), show.data = T)
emtrends(logmar_fd2, pairwise ~ oct_optic_neuritis, var = "fd", infer = T)

logmar_fd3 <-  lme(logmar_min ~ fd*disease_duration + oct_optic_neuritis age + sex  + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_fd3)
intervals(logmar_fd3)
multicollinearity(logmar_fd3)

plot_model(logmar_fd3, type= "emm", terms= c("fd", "disease_duration"))
emtrends(logmar_fd3, pairwise ~ oct_optic_neuritis, var = "fd", infer = T)

models.logmar2 <- list(logmar_gcipl2, logmar_rnfl2, logmar_vad2, logmar_vld2, logmar_fd2)
models.logmar2 <- list(logmar_gcipl2, logmar_rnfl2,logmar_vad2, logmar_vld2, logmar_fd2) %>% 
  set_names('logmar_gcipl', 'logmar_rnfl', 'logmar_vad', 'logmar_vld', 'logmar_fd')

test_performance(models.logmar2)
comparison_logmar<-compare_performance(models.logmar2)
print(comparison_logmar)
compare_performance(models.logmar2) %>% as_tibble() %>% arrange(AIC)
compare_performance(models.logmar2) %>% as_tibble() %>% arrange(BIC)
compare_performance(models.logmar2) %>% as_tibble() %>% arrange(desc(R2_marginal))

compare_performance(models.logmar2) %>% plot
test_performance(logmar_vld2, logmar_vad2, logmar_gcipl2)
#test_vuong(models.logmar2)
general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))

```

## Sensitivity analysis- LogMAR and HE-DMT

```{r}

logmar_vad_het <-  lme(logmar_min ~ vad*oct_optic_neuritis + disease_duration + sex + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_vad_het)
intervals(logmar_vad_het)
check_model(logmar_vad_het)
multicollinearity(logmar_vad_het)

logmar_gcipl_het <-  lme(logmar_min ~ av_6mm*oct_optic_neuritis + disease_duration + sex + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(logmar_gcipl_het)
intervals(logmar_gcipl_het)
check_model(logmar_gcipl_het)
multicollinearity(logmar_gcipl_het)

logmar_rnfl_het <-  lme(logmar_min ~ rnfl_mean_g*oct_optic_neuritis + age + sex  + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(logmar_rnfl_het)
summary(logmar_rnfl_het) 
intervals(logmar_rnfl_het)

```

## Visual outcomes- Sloan 2.5% models

```{r}

#gcipl
sloan2.5_gcipl2 <-  lme(sloan_cont_2_5 ~  av_6mm*oct_optic_neuritis + sex + age + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_gcipl2)
intervals(sloan2.5_gcipl2)
multicollinearity(sloan2.5_gcipl2)
check_model(sloan2.5_gcipl2)

plot_model(sloan2.5_gcipl2, type = 'emm', terms = c('av_6mm', 'oct_optic_neuritis'), show.data = T)
sloan_gcipl<-plot_model(sloan2.5_gcipl2, type = 'emm', terms = c('av_6mm', 'oct_optic_neuritis'), show.data = T)
emtrends(sloan2.5_gcipl2, pairwise ~ oct_optic_neuritis, var = "av_6mm", infer = T)
emtrends(sloan2.5_gcipl2, pairwise ~ oct_optic_neuritis, var = "av_6mm", infer = T) %>% plot()

#prnfl
sloan2.5_rnfl2 <-  lme(sloan_cont_2_5 ~ rnfl_mean_g*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(sloan2.5_rnfl2)
summary(sloan2.5_rnfl2)
intervals(sloan2.5_rnfl2)

plot_model(sloan2.5_rnfl2, type= "emm", terms= c("rnfl_mean_g", "oct_optic_neuritis"))
emtrends(sloan2.5_rnfl2, pairwise ~ oct_optic_neuritis, var = "rnfl_mean_g", infer = T)

general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g)) %>% select(oct_optic_neuritis, rnfl_mean_g)%>%ggpairs()

#vad
sloan2.5_vad2 <-lme(sloan_cont_2_5 ~ vad*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_vad2)
intervals(sloan2.5_vad2)
multicollinearity(sloan2.5_vad2)
check_model(sloan2.5_vad2)

plot_model(sloan2.5_vad2, type = 'emm', terms = c('vad', 'oct_optic_neuritis'), show.data = T)
sloan_vad<-plot_model(sloan2.5_vad2, type = 'emm', terms = c('vad', 'oct_optic_neuritis'), show.data = T)
emtrends(sloan2.5_vad2, pairwise ~ oct_optic_neuritis, var = "vad", infer = T)
emtrends(sloan2.5_vad2, pairwise ~ oct_optic_neuritis, var = "vad", infer = T) %>% plot()

#vld
sloan2.5_vld2 <-lme(sloan_cont_2_5 ~ vld_pct*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_vld2)
intervals(sloan2.5_vld2)
multicollinearity(sloan2.5_vld2)
check_model(sloan2.5_vld2)

plot_model(sloan2.5_vld2, type = 'emm', terms = c('vld_pct', 'oct_optic_neuritis'))
emtrends(sloan2.5_vld2, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T)
emtrends(sloan2.5_vld2, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T) %>% plot()

#fd

sloan2.5_fd2 <-lme(sloan_cont_2_5 ~ fd*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_fd2)
intervals(sloan2.5_fd2)
multicollinearity(sloan2.5_fd2)
check_model(sloan2.5_fd2)

plot_model(sloan2.5_fd2, type = 'emm', terms = c('fd', 'oct_optic_neuritis'), show.data=T)
emtrends(sloan2.5_fd2, pairwise ~ oct_optic_neuritis, var = "fd", infer = T)
emtrends(sloan2.5_fd2, pairwise ~ oct_optic_neuritis, var = "fd", infer = T) %>% plot()


models.sloan2.5_2 <- list(sloan2.5_gcipl2, sloan2.5_rnfl2, sloan2.5_vad2, sloan2.5_vld2, sloan2.5_fd2) %>% 
  set_names('sloan2.5_gcipl', 'sloan2.5_rnfl', 'sloan2.5_vad', 'sloan2.5_vld' ,'sloan2.5_fd')

test_performance(models.sloan2.5_2)
compare_performance(models.sloan2.5_2) %>% as_tibble() %>% arrange(AIC)
compare_performance(models.sloan2.5_2) %>% plot
# test_vuong(models.logmar2)
general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))

aictab(cand.set = models.sloan2.5_2)


models.sloan2.5_2 <- list(sloan2.5_gcipl2, sloan2.5_rnfl2, sloan2.5_vad2, sloan2.5_vld2, sloan2.5_fd2) 
model.names.sloan2.5_2 <-c('gcipl', 'rnfl', 'vad', 'vld', 'fd')
aictab(cand.set = models.sloan2.5_2, model.names.sloan2.5_2)


```

## Sensitivity analysis- Sloan 2.5% and HE-DMT

```{r}

sloan2.5_vad_het <-lme(sloan_cont_2_5 ~ vad*oct_optic_neuritis + age + sex + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_vad_het)
intervals(sloan2.5_vad_het)
multicollinearity(sloan2.5_vad_het)
check_model(sloan2.5_vad_het)

sloan2.5_gcipl_het <-  lme(sloan_cont_2_5 ~  av_6mm*oct_optic_neuritis + sex + age + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(sloan2.5_gcipl_het)
intervals(sloan2.5_gcipl_het)
multicollinearity(sloan2.5_gcipl_het)
check_model(sloan2.5_gcipl_het)

sloan2.5_rnfl_het <-  lme(sloan_cont_2_5 ~ rnfl_mean_g*oct_optic_neuritis + age + sex + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(sloan2.5_rnfl_het)
summary(sloan2.5_rnfl_het)
intervals(sloan2.5_rnfl_het)

```

## Visual outcomes- Ishihara colour models 

```{r}

#gcipl
ishi_gcipl2 <-  lme(lcva_ishihara_plates ~ av_6mm*oct_optic_neuritis+ sex + age + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_gcipl2)
intervals(ishi_gcipl2)
multicollinearity(ishi_gcipl2)
check_model(ishi_gcipl2)

plot_model(ishi_gcipl2, type = 'emm', terms = c('av_6mm', 'oct_optic_neuritis'), show.data=T)
ishihara_gcipl<-plot_model(ishi_gcipl2, type = 'emm', terms = c('av_6mm', 'oct_optic_neuritis'), show.data=T)
emtrends(ishi_gcipl2, pairwise ~ oct_optic_neuritis, var = "av_6mm", infer = T)
emtrends(ishi_gcipl2, pairwise ~ oct_optic_neuritis, var = "av_6mm", infer = T) %>% plot()

#prnfl
ishi_rnfl2 <-  lme(lcva_ishihara_plates ~ rnfl_mean_g*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(ishi_rnfl2)
summary(ishi_rnfl2)
intervals(ishi_rnfl2)

plot_model(ishi_rnfl2, type= "emm", terms= c("rnfl_mean_g", "oct_optic_neuritis"))
emtrends(ishi_rnfl2, pairwise ~ oct_optic_neuritis, var = "rnfl_mean_g", infer = T)
emtrends(ishi_rnfl2, pairwise ~ oct_optic_neuritis, var = "rnfl_mean_g", infer = T) %>% plot()

#vad
ishi_vad2 <-lme(lcva_ishihara_plates ~ vad*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_vad2)
intervals(ishi_vad2)
multicollinearity(ishi_vad2)
check_model(ishi_vad2)

plot_model(ishi_vad2, type = 'emm', terms = c('vad', 'oct_optic_neuritis'), show.data=T)
ishihara_vad<-plot_model(ishi_vad2, type = 'emm', terms = c('vad', 'oct_optic_neuritis'), show.data=T)
emtrends(ishi_vad2, pairwise ~ oct_optic_neuritis, var = "vad", infer = T)
emtrends(ishi_vad2, pairwise ~ oct_optic_neuritis, var = "vad", infer = T) %>% plot()

#vld
ishi_vld2 <-lme(lcva_ishihara_plates ~ vld_pct*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_vld2)
intervals(ishi_vld2)
multicollinearity(ishi_vld2)
check_model(ishi_vld2)

plot_model(ishi_vld2, type = 'emm', terms = c('vld_pct', 'oct_optic_neuritis'))
emtrends(ishi_vld2, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T)
emtrends(ishi_vld2, pairwise ~ oct_optic_neuritis, var = "vld_pct", infer = T) %>% plot()

#fd

ishi_fd2 <-lme(lcva_ishihara_plates ~ fd*oct_optic_neuritis + age + sex + disease_duration + prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_fd2)
intervals(ishi_fd2)
multicollinearity(ishi_fd2)
check_model(ishi_fd2)

plot_model(ishi_fd2, type = 'emm', terms = c('fd', 'oct_optic_neuritis'), show.data = T)
emtrends(ishi_fd2, pairwise ~ oct_optic_neuritis, var = "fd", infer = T)
emtrends(ishi_fd2, pairwise ~ oct_optic_neuritis, var = "fd", infer = T) %>% plot()

models.ishi2 <- list(ishi_gcipl2, ishi_rnfl2, ishi_vad2, ishi_vld2, ishi_fd2) %>% set_names('ishi_gcipl', 'ishi_rnfl', 'ishi_vad', 'ishi_vld', 'ishi_fd')

test_performance(models.ishi2)
compare_performance(models.ishi2) %>% as_tibble() %>% arrange(AIC)
compare_performance(models.ishi2) %>% plot

```

## Sensitivity analysis- Ishihara and HE-DMT

```{r}


ishi_vad_het <-lme(lcva_ishihara_plates ~ vad*oct_optic_neuritis + age + sex + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_vad_het)
intervals(ishi_vad_het)
multicollinearity(ishi_vad_het)
check_model(ishi_vad_het)

AICc(ishi_vad2)
AICc(ishi_vad_het)

ishi_gcipl_het <-  lme(lcva_ishihara_plates ~ av_6mm*oct_optic_neuritis+ sex + age + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
summary(ishi_gcipl_het)
intervals(ishi_gcipl_het)
multicollinearity(ishi_gcipl_het)
check_model(ishi_gcipl_het)

AICc(ishi_gcipl2)
AICc(ishi_gcipl_het)

ishi_rnfl_het <-  lme(lcva_ishihara_plates ~ rnfl_mean_g*oct_optic_neuritis + age + sex + disease_duration + het_prop_dmt, data = {general_P %>% filter(!is.na(m_tort)) %>% filter(!is.na(rnfl_mean_g))}, random=~1|patient_id, na.action=na.omit)
multicollinearity(ishi_rnfl_het)
summary(ishi_rnfl_het)
intervals(ishi_rnfl_het)

AICc(ishi_rnfl2)
AICc(ishi_rnfl_het)

```

## Compare NON eye pairs- difference in colour vision score

```{r}

#Screen out bilateral NON eyes as a separate Df

bilateral_non <- general_P %>%
  group_by(patient_id) %>%                  # replace with your actual subject ID column
  filter(all(oct_optic_neuritis == "no")) %>%  # keeps only subjects where all eyes are non-ON
  ungroup()

bilateral_non%>%distinct(patient_id)
#238 patients

#t-test
ggwithinstats(
  data = bilateral_non,
  x = eye,
  y = lcva_ishihara_plates
)

#Non parametric test
ggwithinstats(
  data = bilateral_non,
  x = eye,
  y = lcva_ishihara_plates,
  type = "nonparametric", ## type of statistical test
  xlab = "Eye", ## label for the x-axis
  ylab = "Ishihara colour vision plates", ## label for the y-axis
  package = "yarrr", ## package from which color palette is to be taken
  palette = "info2", ## choosing a different color palette
  title = "Comparison of Left and Right eyes colour vision",
) + ## modifying the plot further
  ggplot2::scale_y_continuous(
    limits = c(0, 50),
    breaks = seq(from = 0, to = 50, by = 10)
  )


```
